\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-07-05 forwards-and-backwards.tex
minor corrections

\begin{definition}[Generator of a lineage] \label{def:lineage_generator}
    Let $(\varphi_t(x))_{0 \le t \le T}$
    denote the density of a population that solves \eqref{eqn:nonlocal_pde},
    and let $y$ be a point with $\varphi_T(y) > 0$.
    We define $(L_s(y))_{s=0}^T$,
    the ancestral lineage in this population of an individual sampled at $y$ at time $T$,
    to be the position of the unique ancestor of $y$ alive at time $T - s$.
    We define
    $(Q_s)_{s \geq 0}$
    to be the time inhomogeneous semi-group satisfying
    \begin{align*}
        Q_s f(y) := \IE_y[ f(L_s) ] .
    \end{align*}
\end{definition}

@@

As usual,
to make the generator readable, we've written it in concise notation,
omitting the dependencies on location and population density,
which itself changes with time.
When interpreting this,
remember that everything depends on location and density at that location and time --
for instance, ``$r$'' is actually $r(x, \varphi(x))$.

\begin{corollary} \label{cor:lineages_simple}
    In addition to the assumptions of Theorem~\ref{thm:lineages},
    if the dispersal process is isotropic (i.e., $\covq = \sigma^2 I$),
    then
    \begin{equation}
        \Lgen_\varphi f
        =
        \sigma^2 r \gamma
        \left(
            \Delta f
            +
            \left(
                2 \grad \log(\gamma \varphi)
                - \meanq
            \right)
            \cdot \grad f
        \right) .
    \end{equation}
    In other words, 
    a lineage behaves as a diffusion
    that is driven by a Brownian motion 
    run at speed $\sigma^2$ multiplied by the local per-capita production of offspring ($r \gamma$)
    in a potential equal to the local birth output tilted by migration bias
    ($\varphi_s \gamma \exp(-\meanq \cdot x / 2 \sigma^2)$).
\end{corollary}

@@

Comparing \eqref{eqn:pme} to \eqref{eqn:pde},
we see that to set up a limit in which the population density $\varphi$ follows the PME,
we need $r=1$,
$\gamma = \varphi$, and $F = 1 - \varphi$.
Consulting equation~\eqref{eqn:mu_defn},
this implies that $\mu = (1 + 1/\theta) \varphi - 1/\varphi$.
In other words,
establishment is certain
and birth rates increase linearly with population density,
but to compensate, death rates increase slightly faster (also linearly),
as shown in Figure~\ref{fig:pme_waves}.
Alert readers will notice that
for finite $\theta$ death rates may be negative,
and furthermore rates are not bounded;
in practice we choose a suitably small value for $\beta$ and define
\begin{align*}
    \gamma(x, m) &= (1 - \exp(- \beta m)) / \beta \\
    F(x, m) &= \exp(- \beta m) / \beta ,
\end{align*}
and set $\mu(x, m) = \max(0, \gamma - \theta F)$.
Birth and death rates are equal at density $m = 1$,
corresponding to an unscaled density of $N$ individuals per unit area.

@@

%%%%%%%%%%
\paragraph{Fisher--KPP:}
A great deal more attention has been paid to the classical Fisher-KPP equation,
\begin{align} \label{eqn:fkpp}
    v_t = v_{xx} + v (1-v) .
\end{align}
Even though we do not have an explicit formula for the wave shape in this case,
our methods provide information about ancestral lineages.
The equation has non-negative travelling wave solutions of speed $c$ for all $c \geq 2$, 
but started from any compact perturbation of a Heaviside function, the 
solution will converge to the profile $w^F$ with the minimal wavespeed, $c=2$,
\cite{kolmogorov/petrovsky/piscounov:1937,bramson:1983}.
No matter the initial condition,
for any $t>0$ the support of the 
solution will be the whole real line. 
In this case, we must have $r = \gamma = 1$,
and $F(x, m) = 1 - m$ so $\mu(x, m) = 1 - (1-m)/\theta$.
This implies that
the generator of the motion of an ancestral lineage is
\begin{equation} \label{eqn:fkpp_generator}
    \Lgen \phi
    =
    \phi_{xx} + 2 \frac{w_x}{w} \phi_x + 2 \phi_x .
\end{equation}
Near the tip of the wave (for $x$ large), $w^F(x) \sim e^{-x}$,
so \eqref{eqn:fkpp_generator} implies that
the motion of a lineage is close to unbiased Brownian motion.
On the other hand, for $x < 0$ a lineage behaves approximately as
Brownian motion with drift at rate two to the right.
This implies that
ancestral lineages are pushed into the tip of the wave,
and there is no stationary distribution,
so that long-term dynamics of genetic inheritance
depend on the part of the wave not well-approximated by a smooth profile.
These results agree with those of \comment{TODO: citations}.

%%%%%%%%%%
\paragraph{Allen-Cahn:}
Finally, we take the Allen-Cahn equation
\begin{align} \label{eqn:allen_cahn}
    v_t = v_{xx} + v(1-v)(2v-1+s),
\end{align}
for a given $s \in (0,2)$,
which is used to model populations evolving under selection \cite{Sarah}.
This equation also has an explicit travelling wave solution with speed $s$
and shape
\[ w^A(x) = (1+e^{x})^{-1}. \]
The generator for ancestral lineages is, just as in the Fisher-KPP case,
\begin{align*}
    \hat{\mathcal{L}}\phi
    &=
    \phi_{xx}
    + 
    2 \frac{w_x}{w} \phi_x
    +
    s \nabla \phi \\
    \qquad &=
    \phi_{xx}
    -
    2 \frac{e^x}{1+e^x} \phi_x 
    + 
    s \nabla \phi,
\end{align*}
so lineages in the tip are pushed leftwards into the bulk of the wave by a $-2(e^{-x}+1)^{-1}$,
that gets weaker further to the left.
This diffusion has speed measure
$$
    m(d\xi) \propto e^{sx}(1+e^x)^{-2},
$$
which is concentrated around $\log(s/(2-s))$ and decays like $e^{x(s-2)}$ away from this point.
These results agree with those of \cite{etheridge/penington:2020}.

@@

It is natural for applications to wonder about identifiability:
when can the observed quantities like population density
or certain summaries of lineage movement uniquely determine
the underlying demographic parameters?
Consider a deterministic,
continuous population generated by parameters $\gamma$, $\mu$, and $r$
with $meanq = 0$ and $\covq = I$.
Suppose it has a stationary profile $w(x)$, that must satisfy
$$
   r \Delta(\gamma w) + (r \gamma - \mu) w = 0 .
$$
It is easy to see that $w$ does not uniquely specify $\gamma$, $\mu$, and $r$:
let $\beta(x)$ be a smooth, nonnegative function on $\IR^d$,
and let $r'(x, m) = \beta(x) r(x, m)$ and $\mu'(x, m) = \beta(x) \mu(x, m)$
(and, let $\gamma' = \gamma$).
Then the population with parameters $\gamma'$, $\mu'$, and $r'$
has the same stationary profile(s) as the original population.

@@

We use the fact that $\int q_\theta(x,dz)=1$ to rewrite~(\ref{mean measure})
as 
\begin{equation}
\label{eqn:rewritten mean measure}
\begin{split}
    \int\left(
        \int \theta \left( f(z) r(z,\eta)- f(x) r(x,\eta) \right) q_\theta(x,dz)
    \right)
    \gamma(x,\eta)
    \eta(dx)
    \\
    + \int \int f(x) \theta \Big(
        r(x,\eta) \gamma(x,\eta)
        - \mu_\theta(x,\eta)
    \Big) \eta(dx).
\end{split}
\end{equation}
We have defined $\mu_\theta$ so that the second term is simple:
\begin{align*}
    \theta \Big( r(x,\eta) \gamma(x,\eta) - \mu_\theta(x,\eta) \Big)
    = F(x, \eta) .
\end{align*}
Furthermore, recall from Definition~\ref{def:dispersal_generator} that
\begin{align} \label{eqn:near_critical}
    \int \theta \Big(
        r(z,\eta) f(z)
        -
        r(x,\eta) f(x)
    \Big) q_\theta(x,dz) 
    \qquad \stackrel{\theta\to\infty}{\longrightarrow} \qquad  
    \DG \big(r(\cdot,\Xi)f(\cdot)\big)(x) .
\end{align}
To make this more familiar,
remember that if dispersal is simply standard multivariate Gaussian
with mean zero and covariance $\sigma^2 I / \theta$,
then $\DG = \sigma^2 \Delta$, where $\Delta$ denotes the Laplacian.
Equation~\eqref{eqn:rewritten mean measure}) then converges to
\begin{equation} \label{limit of mean measure equation}
\int \gamma(x,\eta)
\DG \big(f(\cdot)r(\cdot,\eta)\big)(x)
\eta(dx)
+
\int f(x)
F(x,\eta)
\eta(dx) .
\end{equation}
This gives the martingale of Theorem~\ref{thm:nonlocal_convergence}.
More generally, we expect everything to work
if the remaining parameters ($r$, $\gamma$, and $F$)
also depended on $N$, but converged as $N \to \infty$.

@@

Although our proof of Theorem~\ref{thm:lineages}
uses the explicit representation in terms of the lookdown process,
informal calculations agree with the main result.
Suppose that we have tracked a lineage to an individual at location $y$ at time $t$.
Looking further back through time, the lineage will move at the time that individual was born,
and will jump to that individual's parents' locations.
Now, the rate at which new individuals are born to parents at $x$ and establish at $y$
is
$$
    \gamma(x, \eta_t) q(x, dy) r(y, \eta_t) .
$$
\comment{TODO: explain this more.}
Happily, we are interested in the case when $\eta$ has a density,
say $\eta_t(dx) = \varphi_t(x) dx$,
so interpreting these rates is not too difficult.
This implies that the probability a randomly chosen individual near $y$
is a new offspring from a parent at $x$ in $[t, t+dt)$ is, informally,
\begin{equation} \label{eqn:informal_rates}
\frac{
    \gamma(x, \eta_t) r(y, \eta_t) \varphi_t(x)
}{
    \varphi_t(y)
} \frac{ q_\theta(x, dy) }{ dy } dx dt .
\end{equation}
Leaving aside questions of whether a lineage can be treated as a randomly chosen individual,
we might then proceed by defining a continuous-time jump process
whose transition rates are given by \eqref{eqn:informal_rates} given $(\varphi_t)_{t=0}^T$.
Let $(L^N_s)_{s=0}^T$ be the location of a lineage that moves according to these jump rates,
back through time (i.e., with the substitution that $s = T-t$).
Then, for the moment writing $q_\theta(x, y)$ for the density of $q_\theta$,
\begin{align} \label{eqn:lineage_generator_Edt}
    \begin{split}
    &\IE[f(L^N_{s+ds}) - f(y) \;|\; L_s = y]
    \\&\qquad 
    =
    ds \int \left(f(x) - f(y)\right)
    \frac{
        \varphi_{T-s}(x) \gamma(x, \eta_{T-s}) r(y, \eta_{T-s})
    }{
        \varphi_{T-s}(y)
    }
    q_\theta(x, y) dx .
    \end{split}
\end{align}
Referring back to Definition~\ref{def:dispersal_generator},
a quick calculation shows that as $N \to \infty$,
\begin{align*}
    \theta \int (f(x) - f(y)) q_\theta(x, y) dx
    &\to
    \DG^* f(x)  - f(x) \DG^* 1 .
\end{align*}
Consequently,
\begin{align*}
    &
    \theta \int (f(x) - f(y)) g(x) q_\theta(x, y) dx \\
    &\qquad =
    \theta \int \left\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \right\} q_\theta(x, y) dx \\
    &\qquad \to
        \DG^*(fg)(y) - f(y) \DG^* g(y) . 
\end{align*}
Applying this to~\eqref{eqn:lineage_generator_Edt} with $g = \varphi_{T-s} \gamma$,
this suggests that the generator of the limiting process is
\begin{align} \label{eqn:heuristic_lineage_generator}
    \Lgen_s f
    &=
    \frac{r}{\varphi_{T-s}}
    \left\{
        \DG^*(\gamma \varphi_{T-s} f) - f \DG^*(\gamma \varphi_{T-s})
    \right\} .
\end{align}
This agrees with Theorem~\ref{thm:lineages}.

@@

\comment{Should we just use $N$ instead of $\lambda$? Change to $N$ but comment that when comparing ot other papers set $\lambda=N$.}

Now we will present a lookdown construction for the general process
(i.e., of Definition \ref{defn:mgale_construction}),
in the spirit of~\cite{kurtz/rodrigues:2011}. 
The general set-up is as follows.
Each individual will be labeled with a ``level'',
which will be a number in $[0, N]$.
We will still encode the process embellished by these levels
as a point measure:
if the $i^\mathrm{th}$ individual's spatial location is $x_i$
and level is $u_i$, then we will write:
$$
    \lp = \sum_i \delta_{x_i, u_i} ,
$$
which is a measure on $\IR^d \times [0, N]$.
Note that each individual contributes mass 1 to the measure,
not $1/N$ as above.
In some sense, these levels do not affect the process:
the pushforward of this measure-valued ``spatial-level process'',
divided by $N$,
will agree with the purely spatial process defined above.
Furthermore, the levels of the individuals present at any spatial location
are conditionally uniform on $[0, N]$.
However, an individual's level in some sense encodes their future reproductive output:
individuals with lower levels tend to live longer, and have more offspring.
For more explanation of the set-up and how this is possible,
see \citet{etheridge/kurtz:2018} and \citet{kurtz/rodrigues:2011}
(and note that our $N$ corresponds to the $\lambda$ of those papers).
Furthermore, the limiting spatial-level process will still be a point measure,
and so we explicitly retain the notion of individuals and lineages through to the 
large-population-density limit.

@@

\comment{
    I've defined $\eta = \lp / N$ to agree with how we define the generator below,
    for which birth/death add/remove something of mass 1, not $1/N$.
}

\comment{TODO: this uses the $\gamma(x, \eta)$ notation, which we haven't settled on.}

Suppose that the initial population is composed of $O(N)$ many particles
with levels uniformly distributed between $[0, N]$.
Suppose the current state of the population is $\lp$,
with spatial projection $\eta$.
An individual at spatial location $x$ with level $u$
produces one juvenile offspring at rate 
$$
2 \left(1 - \frac{u}{N}\right) \gamma(x, \eta) ,
$$
which disperses to a location relative to $x$ drawn from the kernel $q(\cdot)$.
(Averaging over the uniform distribution of the level $u$,
we recover the birth rate $\gamma(x, \eta)$.)
This juvenile -- suppose it's location is $y$ --
either survives, with probability $r(y, \eta)$, or immediately dies.
(As before, ``maturity'' is instantaneous.)

@@

In the lookdown model, levels never cross below 0,
while particles whose levels move above $N$ are regarded as dead
(and do not affect to the population any more).
Therefore, in order to incorporate death,
the level of the individual at location $x$ with level $u$
moves upwards at an additional rate $\mu(x,\eta) u$.
Since levels are uniform,
it is easy to check that if $\mu$ was constant,
this would imply an exponential lifetime for each individual;
see~\cite{etheridge/kurtz:2018}, Section \comment{TODO}
for more general justification.

Putting these together,
the level $u$ of an individual at $x$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \frac{1}{N} u \left(N - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \theta \mu(x,\eta) u .
\end{equation}
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
Recall from equation \eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$,
and so we will have that
\begin{align} \label{eqn:b_limit}
\begin{split}
b(x, \eta)
&=
    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) q(x, dy)
    + F(x, \eta) \\
&\to
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) \qquad \text{as } N \to \infty .
\end{split}
\end{align}

@@

First consider the terms in the generator that come from birth events.
When a birth successfully establishes,
a new level is generated above the parent's level,
and this new level is assigned to either the offspring or the parent.
Since the probability of each is 1/2,
the contribution of birth to the generator is
\begin{align}
f(\lp) \nonumber
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \frac{\theta}{N} \gamma(x, \eta)
    \int_u^N
    \int_{\IR^d}
    \left(
    \frac{1}{2}
    \bigg\{
            g(y, u_1)
        + \frac{ g(y, u) g(x, u_1) }{ g(x, u) }
    \bigg\}
        - 1
    \right)
    r(y, \eta) q(x, dy)
    \\
    \begin{split} \label{eqn:birth_generator}
&=
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 N}
        \int_u^N
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{N}
        \int_u^N \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, u_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align}
Here, $u_1$ is the new level and $y$ is the offspring's location,
and so the two terms in the integral correspond to the two situations:
in the first, we have added an individual at $(y, u_1)$,
while in the second, we replace an individual at $(x, u)$
by one at $(x, u_1)$ and another at $(y, u)$.
We've written it in this form because the two pieces
each naturally converge to separate terms in the limit.

@@

>>>>

\begin{definition}[Generator of a lineage] \label{def:lineage_generator}
    Let $(\varphi_t(x))_{0 \le t \le T}$
    denote the density of a population that solves \eqref{eqn:nonlocal_pde},
    and let $y$ be a point with $\varphi_T(y) > 0$.
    We define $(L_s)_{s=0}^T$,
    the ancestral lineage in this population of an individual sampled at $y$ at time $T$,
    to be the position of the unique ancestor of $y$ alive at time $T - s$.
    We define
    $(Q_s)_{s \geq 0}$
    to be the time inhomogeneous semi-group satisfying
    \begin{align*}
        Q_s f(y) := \IE_y[ f(L_s) ] .
    \end{align*}
\end{definition}

@@

As usual,
to make the generator readable, we've written it in concise notation,
omitting the dependencies on location and population density,
which itself changes with time.
When interpreting this,
remember that everything depends on location and density at that location and time --
for instance, ``$r$'' is actually $r(x, \varphi(x))$ (in the local case),
or $r(x, \smooth{r} \eta(x))$ (in the nonlocal case).

\begin{corollary} \label{cor:lineages_simple}
    In addition to the assumptions of Theorem~\ref{thm:lineages},
    if the dispersal process is isotropic
    (i.e., $\covq = \sigma^2 I$ and $\meanq$ is constant),
    then
    \begin{equation}
        \Lgen_\varphi f
        =
        r \gamma
        \left(
            \sigma^2 \Delta f
            +
            \left(
                2 \sigma^2 \grad \log(\gamma \varphi)
                - \meanq
            \right)
            \cdot \grad f
        \right) .
    \end{equation}
\end{corollary}

In other words, 
the lineage behaves as a diffusion driven by Brownian motion run at speed 
\comment{TODO: make this right:}
If $V_s(x) = \log(\varphi_s(x) \gamma(x, \varphi_s)) - \meanq \cdot x / 2 \sigma^2$,
then the diffusion has drift $\grad V_s$.
If $\varphi$ does not change with time, 
then the stationary distribution is proportional to $\exp(V(x))$.
a lineage behaves as a diffusion
that is driven by a Brownian motion 
run at speed $\sigma^2$ multiplied by the local per-capita production of offspring ($r \gamma$)
in a potential equal to the local birth output tilted by migration bias
($\varphi_s \gamma \exp(-\meanq \cdot x / 2 \sigma^2)$).

@@

Comparing \eqref{eqn:pme} to \eqref{eqn:pde},
we see that to set up a limit in which the population density $\varphi$ follows the PME,
we need $r=1$,
$\gamma = \varphi$, and $F = 1 - \varphi$.
Consulting equation~\eqref{eqn:mu_defn},
this implies that $\mu = (1 + 1/\theta) \varphi - 1/\theta$.
In other words,
establishment is certain
and birth rates increase linearly with population density,
but to compensate, death rates increase slightly faster (also linearly),
as shown in Figure~\ref{fig:pme_waves}.
Alert readers will notice that
for finite $\theta$ death rates may be negative,
and furthermore rates are not bounded;
in practice we choose a suitably small value for $\beta$ and define
\begin{align*}
    \gamma(x, m) &= (1 - \exp(- \beta m)) / \beta \\
    F(x, m) &= \exp(- \beta m) / \beta ,
\end{align*}
and set $\mu(x, m) = \max(0, \gamma - F / \theta)$.
Birth and death rates are equal at density $m = 1$,
corresponding to an unscaled density of $N$ individuals per unit area.

@@

%%%%%%%%%%
\paragraph{Fisher--KPP:}
A great deal more attention has been paid to the classical Fisher-KPP equation,
\begin{align} \label{eqn:fkpp}
    v_t = v_{xx} + v (1-v) .
\end{align}
Even though we do not have an explicit formula for the wave shape in this case,
our methods provide information about ancestral lineages.
The equation has non-negative travelling wave solutions of speed $c$ for all $c \geq 2$, 
but started from any compact perturbation of a Heaviside function, the 
solution will converge to the profile $w^F$ with the minimal wavespeed, $c=2$
\citep{kolmogorov/petrovsky/piscounov:1937,bramson:1983}.
No matter the initial condition,
for any $t>0$ the support of the 
solution will be the whole real line. 
In this case, we must have $r = \gamma = 1$,
and $F(x, m) = 1 - m$ so $\mu(x, m) = 1 + (m-1)/\theta$.
This implies that
the generator of the motion of an ancestral lineage is
\begin{equation} \label{eqn:fkpp_generator}
    \Lgen \phi
    =
    \phi_{xx} + 2 \frac{w_x}{w} \phi_x + 2 \phi_x .
\end{equation}
Near the tip of the wave (for $x$ large), $w^F(x) \sim e^{-x}$,
so \eqref{eqn:fkpp_generator} implies that
the motion of a lineage is close to unbiased Brownian motion.
On the other hand, for $x < 0$ a lineage behaves approximately as
Brownian motion with drift at rate two to the right.
This implies that
ancestral lineages are pushed into the tip of the wave,
and there is no stationary distribution,
so that long-term dynamics of genetic inheritance
depend on the part of the wave not well-approximated by a smooth profile.
These results agree with those of \comment{TODO: citations}.

%%%%%%%%%%
\paragraph{Allen-Cahn:}
Finally, we take the Allen-Cahn equation
\begin{align} \label{eqn:allen_cahn}
    v_t = v_{xx} + v(1-v)(2v-1+s),
\end{align}
for a given $s \in (0,2)$,
which is used to model populations evolving under selection \cite{Sarah}.
This equation also has an explicit travelling wave solution with speed $s$
and shape
\[ w^A(x) = (1+e^{x})^{-1}. \]
The generator for ancestral lineages is, just as in the Fisher-KPP case,
\begin{align*}
    \hat{\mathcal{L}}\phi
    &=
    \phi_{xx}
    + 
    2 \frac{w_x}{w} \phi_x
    +
    s \nabla \phi \\
    \qquad &=
    \phi_{xx}
    -
    2 \frac{e^x}{1+e^x} \phi_x 
    + 
    s \nabla \phi,
\end{align*}
so lineages in the tip are pushed leftwards into the bulk of the wave at a rate $-2(e^{-x}+1)^{-1}$,
that gets weaker further to the left.
This diffusion has speed measure
$$
    m(d\xi) \propto e^{sx}(1+e^x)^{-2},
$$
which is concentrated around $\log(s/(2-s))$ and decays like $e^{x(s-2)}$ away from this point.
These results agree with those of \cite{etheridge/penington:2020}.

@@

It is natural for applications to wonder about identifiability:
when can the observed quantities like population density
or certain summaries of lineage movement uniquely determine
the underlying demographic parameters?
Consider a deterministic,
continuous population generated by parameters $\gamma$, $\mu$, and $r$
with $\meanq = 0$ and $\covq = I$.
Suppose it has a stationary profile $w(x)$, that must satisfy
$$
   r \Delta(\gamma w) + (r \gamma - \mu) w = 0 .
$$
It is easy to see that $w$ does not uniquely specify $\gamma$, $\mu$, and $r$:
let $\beta(x)$ be a smooth, nonnegative function on $\IR^d$,
and let $r'(x, m) = \beta(x) r(x, m)$ and $\mu'(x, m) = \beta(x) \mu(x, m)$
(and, let $\gamma' = \gamma$).
Then the population with parameters $\gamma'$, $\mu'$, and $r'$
has the same stationary profile(s) as the original population.

@@

We use the fact that $\int q_\theta(x,dz)=1$ to rewrite~(\ref{mean measure})
as 
\begin{equation}
\label{eqn:rewritten mean measure}
\begin{split}
    \int\left(
        \int \theta \left( f(z) r(z,\eta)- f(x) r(x,\eta) \right) q_\theta(x,dz)
    \right)
    \gamma(x,\eta)
    \eta(dx)
    \\
    + \int \int f(x) \theta \Big(
        r(x,\eta) \gamma(x,\eta)
        - \mu_\theta(x,\eta)
    \Big) \eta(dx).
\end{split}
\end{equation}
We have defined $\mu_\theta$ so that the second term is simple:
\begin{align*}
    \theta \Big( r(x,\eta) \gamma(x,\eta) - \mu_\theta(x,\eta) \Big)
    = F(x, \eta) .
\end{align*}
Furthermore, recall from Definition~\ref{def:dispersal_generator} that
\begin{align} \label{eqn:near_critical}
    \int \theta \Big(
        r(z,\eta) f(z)
        -
        r(x,\eta) f(x)
    \Big) q_\theta(x,dz) 
    \qquad \stackrel{\theta\to\infty}{\longrightarrow} \qquad  
    \DG \big(r(\cdot,\Xi)f(\cdot)\big)(x) .
\end{align}
To make this more familiar,
remember that if dispersal is simply standard multivariate Gaussian
with mean zero and covariance $\sigma^2 I / \theta$,
then $\DG = \sigma^2 \Delta$, where $\Delta$ denotes the Laplacian.

In summary, equation~\eqref{eqn:rewritten mean measure} converges to
\begin{equation} \label{limit of mean measure equation}
\int \gamma(x,\eta)
\DG \big(f(\cdot)r(\cdot,\eta)\big)(x)
\eta(dx)
+
\int f(x)
F(x,\eta)
\eta(dx) .
\end{equation}
This gives the martingale of Theorem~\ref{thm:nonlocal_convergence}.
More generally, we expect everything to work
if the remaining parameters ($r$, $\gamma$, and $F$)
also depended on $N$, but converged as $N \to \infty$.

@@

Although our proof of Theorem~\ref{thm:lineages}
uses the explicit representation in terms of the lookdown process,
informal calculations agree with the main result.
Suppose that we have tracked a lineage back to an individual at location $y$ at time $t$.
Looking further back through time, the lineage will move at the time that individual was born,
and will jump to that individual's parents' locations.
Now, the rate at which new individuals are born to parents at $x$ and establish at $y$
is
$$
    \gamma(x, \eta_t) q(x, dy) r(y, \eta_t) .
$$
\comment{TODO: explain this more.}
Happily, we are interested in the case when $\eta$ has a density,
say $\eta_t(dx) = \varphi_t(x) dx$,
so interpreting these rates is not too difficult.
This implies that the probability a randomly chosen individual near $y$
is a new offspring from a parent at $x$ in $[t, t+dt)$ is, informally,
\begin{equation} \label{eqn:informal_rates}
\frac{
    \gamma(x, \eta_t) r(y, \eta_t) \varphi_t(x)
}{
    \varphi_t(y)
} \frac{ q_\theta(x, dy) }{ dy } dx dt .
\end{equation}
Leaving aside questions of whether a lineage can be treated as a randomly chosen individual,
we might then proceed by defining a continuous-time jump process
whose transition rates are given by \eqref{eqn:informal_rates} given $(\varphi_t)_{t=0}^T$.
Let $(L^N_s)_{s=0}^T$ be the location of a lineage that moves according to these jump rates,
back through time (i.e., with the substitution that $s = T-t$).
Then, for the moment writing $q_\theta(x, y)$ for the density of $q_\theta$,
\begin{align} \label{eqn:lineage_generator_Edt}
    \begin{split}
    &\IE[f(L^N_{s+ds}) - f(y) \;|\; L_s = y]
    \\&\qquad 
    =
    ds \int \left(f(x) - f(y)\right)
    \frac{
        \varphi_{T-s}(x) \gamma(x, \eta_{T-s}) r(y, \eta_{T-s})
    }{
        \varphi_{T-s}(y)
    }
    q_\theta(x, y) dx .
    \end{split}
\end{align}
Referring back to Definition~\ref{def:dispersal_generator},
a quick calculation shows that as $N \to \infty$,
\begin{align*}
    \theta \int (f(x) - f(y)) q_\theta(x, y) dx
    &\to
    \DG^* f(x)  - f(x) \DG^* 1 .
\end{align*}
Consequently,
\begin{align*}
    &
    \theta \int (f(x) - f(y)) g(x) q_\theta(x, y) dx \\
    &\qquad =
    \theta \int \left\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \right\} q_\theta(x, y) dx \\
    &\qquad \to
        \DG^*(fg)(y) - f(y) \DG^* g(y) . 
\end{align*}
Applying this to~\eqref{eqn:lineage_generator_Edt} with $g = \varphi_{T-s} \gamma$,
this suggests that the generator of the limiting process is
\begin{align} \label{eqn:heuristic_lineage_generator}
    \Lgen_s f
    &=
    \frac{r}{\varphi_{T-s}}
    \left\{
        \DG^*(\gamma \varphi_{T-s} f) - f \DG^*(\gamma \varphi_{T-s})
    \right\} .
\end{align}
This agrees with Theorem~\ref{thm:lineages}.

@@

Now we will present a lookdown construction for the general process
(i.e., of Definition \ref{defn:mgale_construction}),
in the spirit of~\cite{kurtz/rodrigues:2011}. 
The general set-up is as follows.
Each individual will be labeled with a ``level'',
which will be a number in $[0, N]$.
We will still encode the process embellished by these levels
as a point measure:
if the $i^\mathrm{th}$ individual's spatial location is $x_i$
and level is $u_i$, then we will write:
$$
    \lp = \sum_i \delta_{x_i, u_i} ,
$$
which is a measure on $\IR^d \times [0, N]$.
Note that each individual contributes mass 1 to the measure,
not $1/N$ as above.
In some sense, these levels do not affect the process:
the pushforward of this measure-valued ``spatial-level process'',
divided by $N$,
will agree with the purely spatial process defined above.
Furthermore, the levels of the individuals present at any spatial location
will be conditionally uniform on $[0, N]$.
However, an individual's level in some sense encodes their future reproductive output:
individuals with lower levels tend to live longer, and have more offspring.
For more explanation of the set-up and how this is possible,
see \citet{etheridge/kurtz:2018} and \citet{kurtz/rodrigues:2011}
(and note that our $N$ corresponds to the $\lambda$ of those papers).
Furthermore, the limiting spatial-level process will still be a point measure,
and so we explicitly retain the notion of individuals and lineages through to the 
large-population-density limit.

@@

Suppose that the initial population is composed of $O(N)$ many particles
with levels uniformly distributed between $[0, N]$.
Suppose the current state of the population is $\lp$,
with spatial projection $\eta$.
An individual at spatial location $x$ with level $u$
produces one juvenile offspring at rate 
$$
2 \left(1 - \frac{u}{N}\right) \gamma(x, \eta) ,
$$
which disperses to a location relative to $x$ drawn from the kernel $q(x, \cdot)$.
(Averaging over the uniform distribution of the level $u$,
we recover the birth rate $\gamma(x, \eta)$.)
This juvenile -- suppose it's location is $y$ --
either survives, with probability $r(y, \eta)$, or immediately dies.
(As before, ``maturity'' is instantaneous.)

@@

In the lookdown model, levels never cross below 0,
while particles whose levels move above $N$ are regarded as dead
(and do not affect to the population any more).
Therefore, in order to incorporate death,
the level of the individual at location $x$ with level $u$
moves upwards at an additional rate $\mu(x,\eta) u$.
Since levels are uniform,
it is easy to check that if $\mu$ were constant,
this would imply an exponential lifetime for each individual;
see~\cite{etheridge/kurtz:2018}, Section \comment{TODO}
for more general justification.

Putting these together,
the level $u$ of an individual at $x$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \frac{1}{N} u \left(N - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \mu(x,\eta) u .
\end{equation}
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
Recall from equation \eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$,
and so we will have that
\begin{align} \label{eqn:b_limit}
\begin{split}
b(x, \eta)
&=
    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) q(x, dy)
    + F(x, \eta) \\
&\to
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) \qquad \text{as } \theta \to \infty .
\end{split}
\end{align}

@@

First consider the terms in the generator that come from birth events.
When a birth successfully establishes,
a new level is generated above the parent's level,
and this new level is assigned to either the offspring or the parent.
Since the probability of each is 1/2,
the contribution of birth to the generator maps $f(\lp)$ to
\begin{align}
    &
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \frac{\theta}{N} \gamma(x, \eta)
    \int_u^N
    \int_{\IR^d}
    \left(
    \frac{1}{2}
    \bigg\{
            g(y, u_1)
        + \frac{ g(y, u) g(x, u_1) }{ g(x, u) }
    \bigg\}
        - 1
    \right)
    r(y, \eta) q(x, dy)
    du_1
    \\
    \begin{split} \label{eqn:birth_generator}
&=
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 N}
        \int_u^N
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{N}
        \int_u^N \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, u_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
        du_1
    \bigg\}
    .
    \end{split}
\end{align}
Here, $u_1$ is the new level and $y$ is the offspring's location,
and so the two terms in the integral correspond to the two situations:
in the first, we have added an individual at $(y, u_1)$,
while in the second, we replace an individual at $(x, u)$
by one at $(x, u_1)$ and another at $(y, u)$.
We've rewritten it in this form because the two pieces
each naturally converge to separate terms in the limit.

@@

\comment{TODO: explain a bit more about the infinitely many points in the $v$ direction.}
